\ieeetitle

### 2016-12-29 report/introduction.tex
pass on thesis

Creating an algorithm providing scalability, integrity and validity, along with a total ordering for the events through all peers in a distributed system has been one the hot topics in distributed systems research for many years. One of the recently designed algorithms on this topic is \epto \autocite{matos2015epto}. \epto is an algorithm that claims to provide integrity, validity and total order in a large-scale distributed system. In addition, \epto is designed to work without a global clock, removing the need to synchronize clocks precisely on every peer and thus is well-suited for dynamic large-scale distributed systems.

@@

There are many other algorithms for disseminating and ordering events in a distributed system. There are some deterministic algorithms, which guarantee total order, agreement or other strong properties. Unfortunately, these types of algorithms are not scalable enough to be used in a large-scale distributed system \autocites[]{defago2004total}[]{lamport1978time}.

@@

explicitly maintain a group and have access to a failure detector \autocites[]{chandra1996weakest}[]{chandra1996unreliable}. Due to faults and churn in large-scale distributed systems, the failure detector turns into a bottleneck for the system and thus limits the algorithm's scalability.

@@

As an alternative to deterministic algorithms, there are probabilistic algorithms, focusing on scalability and resiliency against failures using a probabilistic dissemination approach \autocites []{birman1999bimodal}[]{carvalho2007emergent}[]{demers1987epidemic}[]{eugster2003lightweight}[]{felber2002probabilistic}[]{hayden1996probabilistic}[]{kim2004gossip}[]{Koldehofe02simplegossiping}. These algorithms guarantee the dissemination of events in the system with high probability. This way, there is no need for failure detectors and redundant traffic, making these algorithms highly scalable. As these algorithms focus on reliability of dissemination, they often have to ignore other properties such as total ordering.

@@

This is where \epto shines. \epto, by mixing these two types of algorithms, provides total order along with scalability, validity and integrity. 
\epto consists of two distinct parts. The first part is probabilistic dissemination. \epto guarantees that all peers will receive an event with arbitrarily high probability. The second part of \epto is deterministic ordering. Once peers have received all events, they deterministically order them using the events timestamp, and in case of a tie use the broadcaster ID of the events.

@@

Using the balls-and-bins approach we model processes as bins and events as balls and calculate how many balls need to be \textit{thrown} such that every bin contains at least one ball with arbitrarily high probability. With this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore the number of messages sent in the system is low and uniform over all processes. Thanks to these approaches, \epto becomes highly scalable and resilient, while still providing total order.

@@

The next sections are dedicated to present our benchmarking solution and to present the results obtained. \autoref{sec:ordering} presents the different kinds of ordering. \autoref{sec:definitions} defines the terms used in the paper. \autoref{sec:epto} presents \epto and the architecture of our implementation. \autoref{sec:evaluation} shows and explains the results obtained as well as the limitations our project faced. We then present possible future work in \autoref{sec:future-work}. We finally conclude in \autoref{sec:conclusion}.

>>>>

Creating an algorithm providing scalability, integrity and validity, along with a total ordering of the events through all peers in a distributed system has been one the hot topics in distributed systems research for many years. One of the recently designed algorithms on this topic is \epto \autocite{matos2015epto}. \epto is an algorithm that claims to provide integrity, validity and total order in a large-scale distributed system. In addition, \epto is designed to work without a global clock, removing the need to synchronize clocks precisely on every peer and thus is well-suited for dynamic large-scale distributed systems.

@@

There are many other algorithms for disseminating and ordering events in a distributed system. There are some deterministic algorithms, which guarantee total order, agreement or other strong properties. Unfortunately, these types of algorithms are not scalable enough to be used in large-scale distributed systems \autocites[]{defago2004total}[]{lamport1978time}.

@@

explicitly maintain a group and have access to a failure detector \autocites[]{chandra1996weakest}[]{chandra1996unreliable}. Due to faults and churn in large-scale distributed systems, the failure detector turns into a bottleneck for the system and thus limits the algorithm's scalability. \hm{The first three paragraphs use the word system a lot. You could rework this a little bit.}

@@

As an alternative to deterministic algorithms, there are probabilistic algorithms, focusing on scalability and resiliency against failures using a probabilistic dissemination approach \hm{Write [6-13] instead.} \autocites []{birman1999bimodal}[]{carvalho2007emergent}[]{demers1987epidemic}[]{eugster2003lightweight}[]{felber2002probabilistic}[]{hayden1996probabilistic}[]{kim2004gossip}[]{Koldehofe02simplegossiping}. These algorithms guarantee the dissemination of events in the system with high probability. This way, there is no need for failure detectors and redundant traffic, making these algorithms highly scalable. As these algorithms focus on reliability of dissemination, they often have to ignore other properties such as total ordering.

@@

This is where \epto shines. \hm{What does ``where'' refer to? Not clear.} \epto, by mixing these two types of algorithms, \hm{Not sure I agree here. What two types? Not clear. Epto is not probabilistic and deterministic. Please rephrase.} provides total order along with scalability, validity and integrity. 
\epto consists of two distinct parts. The first part is probabilistic dissemination. \epto guarantees that all peers will receive an event with arbitrarily high probability. The second part of \epto is deterministic ordering. Once peers have received all events, they deterministically order them using the events timestamp, and in case of a tie use the broadcaster ID of the events. \hm{Write component instead of part. Fix before and after.}

@@

Using the balls-and-bins approach we model processes as bins and events as balls and calculate how many balls need to be \textit{thrown} such that every bin contains at least one ball with arbitrarily high probability. With this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore the number of messages sent in the system is low and uniform over all processes. Thanks to these approaches, \epto is highly resilient with  imperfect networks and highly scalable as the networks grow, while always providing total order. \par

@@

The next sections present our benchmarking solution and the results obtained. \autoref{sec:ordering} presents the different kinds of ordering. \autoref{sec:definitions} defines the terms used in the paper. \autoref{sec:epto} presents \epto and the architecture of our implementation. \autoref{sec:evaluation} shows and explains the results obtained as well as the limitations our project faced. We then present possible future work in \autoref{sec:future-work}. We finally conclude in \autoref{sec:conclusion}. \hm{Write Section X instead of section X. Scan all document.}

### 2016-12-29 report/main.tex
Continue work on thesis
WIP report
pass on thesis

\sisetup{
	round-mode          = places,
	round-precision     = 2,
}
\DeclareSIUnit\mbps{MB/s}

\pgfplotscreateplotcyclelist{my colors}{%
solid, color=black\\%
densely dotted, color=blue\\%
densely dashed, color=red\\%
}
\makeatother
%% Environment for comments: Set the boolean to false to produce a comment-free version.
\newboolean{showcomments}
\setboolean{showcomments}{true}
\ifthenelse{\boolean{showcomments}}
{ \newcommand{\mynote}[3]{
		\fbox{\bfseries\sffamily\scriptsize#1}
		{\small$\blacktriangleright$\textsf{\textit{\color{#3}{#2}}}$\blacktriangleleft$}}}
{ \newcommand{\mynote}[3]{}}
% One command per author:
\newcommand{\jt}[1]{\mynote{Jocelyn}{#1}{blue}}

@@

\pgfplotsset{
	compat=1.14,
	tick label style={font=\small},
	label style={font=\small},
	legend style={font=\small},
	/pgfplots/short line legend/.style={
		legend image code/.code={
			\draw[mark repeat=2,mark phase=2,##1]
			plot coordinates {
				(0cm,0cm)
				(0.2cm,0cm)
				(0.4cm,0cm)
			};%
		}
	},
	/pgfplots/xtick pos=left,
}

@@

			% === Logos ==============================================
			\begin{figure}[htp]
				\centering
				\subfloat{\includegraphics[scale=0.60]{logos/UNI_Bern.pdf}}\hfill
				\subfloat{\includegraphics[scale=0.54]{logos/UNI_Neuchatel.pdf}}\hfill
				\subfloat{\includegraphics[scale=0.81]{logos/UNI_Fribourg.pdf}}
			\end{figure}
			% === // Logos ===========================================

@@

\input{abstract.tex}
\input{introduction.tex}
\input{ordering.tex}
\input{definitions.tex}
\input{framework.tex}
\input{evaluation.tex}
\input{future-work.tex}
\input{conclusion.tex}

\printbibliography
\end{document}

>>>>

\sisetup{
	round-mode          = places,
	round-precision     = 2,
}
\DeclareSIUnit\mbps{MB/s}
\DeclareSIUnit\gbps{Gb/s}

\pgfplotscreateplotcyclelist{my colors}{%
solid, color=black\\%
densely dotted, color=blue\\%
densely dashed, color=red\\%
}
\makeatother
%% Environment for comments: Set the boolean to false to produce a comment-free version.
\newboolean{showcomments}
\setboolean{showcomments}{true}
\ifthenelse{\boolean{showcomments}}
{ \newcommand{\mynote}[3]{
		\fbox{\bfseries\sffamily\scriptsize#1}
		{\small$\blacktriangleright$\textsf{\textit{\color{#3}{#2}}}$\blacktriangleleft$}}}
{ \newcommand{\mynote}[3]{}}
% One command per author:
\newcommand{\jt}[1]{\mynote{Jocelyn}{#1}{blue}}
\newcommand{\hm}[1]{\mynote{Hugues}{#1}{orange}}

@@

\pgfplotsset{
	% compat=1.14,
	tick label style={font=\small},
	label style={font=\small},
	legend style={font=\small},
	/pgfplots/short line legend/.style={
		legend image code/.code={
			\draw[mark repeat=2,mark phase=2,##1]
			plot coordinates {
				(0cm,0cm)
				(0.2cm,0cm)
				(0.4cm,0cm)
			};%
		}
	},
	/pgfplots/xtick pos=left,
}

@@

			% === Logos ==============================================
			\begin{figure}[htp]
				\centering
				\subfloat{\includegraphics[scale=0.60]{logos/UNI_Bern.pdf}}\hfill
				\subfloat{\includegraphics[scale=0.54]{logos/UNI_Neuchatel.pdf}}\hfill
				\subfloat{\includegraphics[scale=0.81]{logos/UNI_Fribourg.pdf}}
			\end{figure}
			% === // Logos ==========thun=================================

@@

\input{abstract.tex}
\input{introduction.tex}
\input{ordering.tex}
\input{definitions.tex}
\input{framework.tex}
\input{evaluation.tex}
\input{future-work.tex}
\input{conclusion.tex}
\printbibliography
\end{document}

### 2016-12-29 report/ordering.tex
pass on thesis

\section{Ordering Algorithms}
\label{sec:ordering}
\jt{Do we still want this section?}
Distributed systems and centralized systems alike, need to preserve the temporal order of events produced by concurrent processes in the system. When there are separated processes that can only communicate through messages, you cannot easily order these messages.
Therefore we need ordering algorithms to overcome this problem.
\par
We have two types of ordering algorithms \autocite{lamport1978time}: the partial order algorithms and the total order algorithms.
\subsection{Partial Order Algorithms}
Assuming S is partially ordered under $\leq$, then the following statements hold for all a, b and c in S:
\begin{itemize}
	\item Reflexivity: $a \leq a$ for all $a \in S$.
	\item Antisymmetry: $a \leq b$ and $b \leq a$ implies $a=b$ .
	\item Transitivity: $a \leq b$  and $b \leq c$  implies $a \leq c$.
\end{itemize}

\subsection{Total Order Algorithms}
A totally ordered set of events is a partially ordered set which satisfies one additional property:
\begin{itemize}
	\item Totality (trichotomy law): For any $a, b \in S$, either $a \leq b$  or $b \leq a$.
\end{itemize}
\par
In other words, total order is an ordering that defines the exact order of every event in the system. On the other hand, partial ordering only defines the order between certain key events that depend on each other. Partial order can be useful since it is less costly to implement. However, in some cases the order of all events is important, for example when we need to know exactly which operations have been invoked in which order. We then have to use total order, otherwise we can end up in an inconsistent state.

>>>>

\section{Ordering Algorithms}
\label{sec:ordering}
\jt{Do we still want this section?}\hm{Absolutely!}
Distributed systems and centralized systems alike, need to preserve the temporal order of events produced by concurrent processes in the system. When there are separated processes that can only communicate through messages, you cannot easily order these messages.
Therefore we need ordering algorithms to overcome this problem.
\par
We have two types of ordering algorithms \autocite{lamport1978time}: the partial order algorithms and the total order algorithms.
\subsection{Partial Order Algorithms}
Assuming S is partially ordered under $\leq$, then the following statements hold for all a, b and c in S:
\begin{itemize}
	\item Reflexivity: $a \leq a$ for all $a \in S$.
	\item Antisymmetry: $a \leq b$ and $b \leq a$ implies $a=b$ .
	\item Transitivity: $a \leq b$  and $b \leq c$  implies $a \leq c$.
\end{itemize}

\subsection{Total Order Algorithms}
A totally ordered set of events is a partially ordered set which satisfies one additional property:
\begin{itemize}
	\item Totality (trichotomy law): For any $a, b \in S$, either $a \leq b$  or $b \leq a$.
\end{itemize}
\par
In other words, total order is an ordering that defines the exact order of every event in the system. On the other hand, partial ordering only defines the order between certain key events that depend on each other. Partial order can be useful since it is less costly to implement. However, in some cases the order of all events is important, for example when we need to know exactly which operations have been invoked in which order. We then have to use total order, otherwise we can end up in an inconsistent state. \hm{It would be enlightening to be more precise here and add a large-scale distributed application for which total order is important.}
